\documentclass[main.tex]{subfiles}

\begin{document}
%\begin{titlepage}
\thispagestyle{empty}
%%%%%%%%%
% TITLE %
%%%%%%%%%
{
\noindent
\Large
\bfseries
Topological symmetry-restored phase of gravity
\bigskip
\bigskip
}

%%%%%%%%%%%
% AUTHORS %
%%%%%%%%%%%
%\begin{addmargin}[20mm]{0mm}
{
\noindent
\bfseries
Guilherme~Sadovski${}^{1}${\normalfont~and}
Rodrigo~F.~Sobreiro${}^{2}$
\bigskip
}

%%%%%%%%%%%%%%%
% AFFILIATION %
%%%%%%%%%%%%%%%
{
\noindent
\footnotesize
${}^{1}$HENU --- Henan University, Institute of Contemporary Mathematics, School of Mathematics and Statistics, Kaifeng, Henan 475004, P. R. China. \\
${}^{2}$UFF --- Universidade Federal Fluminense, Instituto de F{\'{\i}}sica, Campus da Praia Vermelha, Av. Litor{\^{a}}nea s/n, 24210--346, Niter{\'{o}}i, RJ, Brazil.
\bigskip
\bigskip
}

%%%%%%%%%%
% EMAILS %
%%%%%%%%%%
{
\noindent
\footnotesize
\rmfamily
e-mails:~\href{
  mailto:gsadovski@henu.edu.cn}{
  gsadovski@henu.edu.cn}{\normalfont~and}
\href{
  mailto:rodrigo\_sobreiro@id.uff.br
}{
  rodrigo\_sobreiro@id.uff.br}{\normalfont.}
\bigskip
}

%%%%%%%%%%%%
% ABSTRACT %
%%%%%%%%%%%%
{
\noindent
\bfseries
Abstract.
\normalfont{}
In this work, we proposed a topological quantum field theory phase for four-dimensional gravity. We show it is able to generate not only General Relativity, but the whole family of Lovelock-Cartan theories of gravity. This is accomplished due to the existence of a topological symmetry that, when explicitly broken via the introduction of a mass scale, releases the local degrees of freedom of gravity in the bulk. Additionally, we evaluate its Ward identities, counterterms, and prove its quantum stability, to all orders in perturbation theory, using the algebraic renormalization technique, and an extended notion of the (anti-){}self-dual Landau gauge conditions.
\bigskip
}
%\end{addmargin}

%%%%%%%
% TOC %
%%%%%%%
{
\noindent
\rule{\textwidth}{1pt}
\tableofcontents
\smallskip
\noindent
\rule{\linewidth}{1pt}
}
%\end{titlepage}
\end{document}
